Ancestry linkage in hybrid zones

### 2016-02-12 drafts/hybrid_zone.tex
made some edits to the intro

Within a genome, the mixing of ancestries is brought about by the the process of recombination that takes place every generation. The expected distribution of lengths of unbroken ancestry have been previously described under neutral conditions \cite[e.g.][]{Gravel2012,Sedghifar2015}.  
Here, we investigate expected patterns of co-ancestry \yb{ --- that is, the genomic scale at which ancestry strongly covaries between sites --- } across a hybrid zone in the presence of selection. 
Selection against migrant \yb{alleles} in a hybrid zone results in reduced frequencies of foreign ancestry surrounding targets of selection. 
The effect on patterns of ancestral block lengths is, however, less well understood. 
Because of rapid removal from the population, we expect to observe \yb{a deficit of} short blocks of foreign ancestry surrounding the selected locus, with this effect becoming more pronounced further away from the center of a hybrid zone.

Some words about how new genomic data will allow us to focus on tracts and not just individual loci.

While population genetic models of hybrid zones have been extensively explored, these have typically focused on the equilibrium states of ancestry clines \cite{Barton1979a,Barton1986}. 
These models predict cline width to be narrowest at loci close to selected loci, and cline width at less tightly linked loci to be wider and more heavily influenced by hybrid zone age and rates of gene flow. 
\yb{The focus on cline width provide an expectation for how selection shapes the covariation in ancestry at a single locus over \emph{geographic space}, our aim is to provide an expectation for how selection shapes the covariation in ancestry at loci over \emph{genomic space.}}  
As denser genotype data becomes more readily available, genome-wide patterns of linkage can be utilized in the analysis of gene hybrid zones.

% Genome-wide covariance in ancestry can provide information about the number of surviving recombination events that have taken place between ancestral genotypes and therefore provides information about the age of the hybrid zone, and the strength of selection at linked loci.  %YB Genome-wide sounds wrong to me. is this ok?
Covariance in ancestry across a genome can provide information about the number of surviving recombination events that have taken place between ancestral genotypes and therefore provides information about the age of the hybrid zone, and the strength of selection at linked loci. 
%Our goal here is  to understand how strength of selection, density of selected sites, recombination rate and hybrid zone age interact to produce observed patterns of ancestry in present day populations.   % YB I don't think we are actually asking about density of selected sites, are we? 
Our goal here is  to understand how strength of \yb{selection, recombination} rate and hybrid zone age interact to produce observed patterns of ancestry in present day populations.  
This may be especially important for an understanding and interpretation of of non-equilibrium patterns of diversity in hybrid zones, as well as understanding the consequences of selection on the extent of mixing of ancestral genomes and, consequently, the maintenance of species distinctions. In the extreme case of zero fitness in first-generation hybrids, for example, paternal genomes will never mix and the stable hybrid zone will resemble a recently formed one. If hybrids have non-zero fitness, and some amount of introgression can occur, and patterns of introgression can provide information about the strength of selection, extent of hybridization, and timing of secondary contact. In this latter case, it may be of additional interest to establish whether the hybrid zone is in its equilibrium state.

Expectations:
Conditional  on having identity B on the wrong side, far away from the center of the zone, the length of unbroken ancestry surrounding the selected locus should be relatively long. This is because an unfit haplotype is likely to be a recent migrant from the other side of the hybrid zone, and therefore will not have experienced many recombination events.

@@

\plr{or, perhaps for introduction?}
In addition to our finding that ancestry surrounding selected sites was quite long, our simulations revealed a consistent and unexpected finding. We found, on a given side of the hybrid zone, a transition from the locally adaptive ancestry at the selected site to the alternative ancestry at a closely linked site was short-lived. Put differently, blocks of `wrong' ancestry that do not overlap the selected site but are closely linked are exceptionally short. A simple intuitive explanation for this pattern suggests itself - loci physically linked to the selected sites can only survive on the wrong side of the hybrid zone for a reasonable amount of time if they recombine away from their ancestry at the locally adaptive site. Because it is then tightly linked to the alternate ancestry at the selected site, this ancestry chunk will large recombine with the alternate ancestry, and therefore decay rapidly. This pattern means that we can further refine the signal of block lengths around a selected locus, by measuring the ratio $\frac{2\sum{l_B(m_i)}}{\sum{l_A(m_i+)+l_A(m_i-)}}$ of mean block lengths surrounding a site with the mean length of adjacent blocks (Figures~\ref{Supp:ratioBlockAdjacent},~\ref{Supp:ratioBlockAdjacentHeatmap})

@@

\begin{figure}
\includegraphics{figs/number_of_ancestors_tau1000}
\caption{The number of individuals of ancestry B per number of ancestors from $\tau=1000$ represented across geographic space. Each line }\label{Fig:family_size}
\end{figure}

>>>>

Within a genome, the mixing of ancestries is brought about by the the process of recombination that takes place every generation. The expected distribution of lengths of unbroken ancestry have been previously described under neutral conditions \cite[e.g.][]{Gravel2012,Sedghifar2015}.  
Here, we investigate expected patterns of co-ancestry \yb{ --- that is, the genomic scale at which ancestry strongly covaries between sites --- } across a hybrid zone in the presence of selection. 
Selection against migrant \yb{alleles} in a hybrid zone results in reduced frequencies of foreign ancestry surrounding targets of selection. 
The effect on patterns of ancestral block lengths is, however, less well understood. 
Because of rapid removal from the population, we expect to observe \yb{a deficit of} short blocks of foreign ancestry surrounding the selected locus, with this effect becoming more pronounced further away from the center of a hybrid zone. \alisa{While empirical patterns of block length in hybrid zones have not been well studied, the availability of dense genotype data in combination with linkage information pave a way to utilize such patterns to better understand the processes operating in hybrid zones. }

While population genetic models of hybrid zones have been extensively explored, these have typically focused on the equilibrium states of ancestry clines \cite{Barton1979a,Barton1986}. 
These models predict cline width to be narrowest at loci close to selected loci, and cline width at less tightly linked loci to be wider and more heavily influenced by hybrid zone age and rates of gene flow. 
\yb{The focus on cline width provide an expectation for how selection shapes the covariation in ancestry at a single locus over \emph{geographic space}, our aim is to provide an expectation for how selection shapes the covariation in ancestry at loci over \emph{genomic space.}}  
%As denser genotype data becomes more readily available, genome-wide patterns of linkage can be utilized in the analysis of gene hybrid zones.

% Genome-wide covariance in ancestry can provide information about the number of surviving recombination events that have taken place between ancestral genotypes and therefore provides information about the age of the hybrid zone, and the strength of selection at linked loci.  %YB Genome-wide sounds wrong to me. is this ok?
Covariance in ancestry across a genome can provide information about the number of surviving recombination events that have taken place between ancestral genotypes and therefore provides information about the age of the hybrid zone and the strength of selection at linked loci. 
%Our goal here is  to understand how strength of selection, density of selected sites, recombination rate and hybrid zone age interact to produce observed patterns of ancestry in present day populations.   % YB I don't think we are actually asking about density of selected sites, are we? 
Our goal here is  to understand how strength of \yb{selection, recombination} rate, and hybrid zone age interact to produce observed patterns of ancestry in present day populations.  
This may be especially important for our understanding and interpretation of of non-equilibrium patterns of diversity in hybrid zones, as well as understanding the consequences of selection on the extent of mixing of ancestral genomes and, consequently, the maintenance of species distinctions. In the extreme case of zero fitness in first-generation hybrids, for example, paternal genomes will never mix and the stable hybrid zone will resemble a recently formed one. If hybrids have non-zero fitness, and some amount of introgression can occur, and patterns of introgression can provide information about the strength of selection, extent of hybridization, and timing of secondary contact. In this latter case, it is also of interest to identify putative targets of selection, as these will provide insights into the biological mechanisms of speciation. 
%In this latter case, it may be of additional interest to establish whether the hybrid zone is in its equilibrium state.

Our naive expectation under a basic underdominance model is that, conditional on having the ancestry that is at lower frequency (that is, being on the `wrong' side of the hybrid zone), the length of unbroken ancestry surrounding the selected locus will be relatively long when far away from the zone center. This is because an unfit haplotype is likely to be a recent migrant from the other side of the hybrid zone, and therefore will not have experienced many recombination events. Using a combination of predictions from theory and simulated hybrid zones, we present a model of genome-wide patterns coancestry as they relate to hybrid zone age, genetic distance from selected locus and geographic distance from hybrid zone center.

\alisa{Moved some points here to the discussion.}

@@

\plr{or, perhaps for introduction?}
In addition to our finding that ancestry surrounding selected sites was quite long, our simulations revealed a consistent and unexpected finding. We found, on a given side of the hybrid zone, a transition from the locally adaptive ancestry at the selected site to the alternative ancestry at a closely linked site was short-lived (Figure~\ref{Supp:adjacentBlocks}). Put differently, blocks of `wrong' ancestry that do not overlap the selected site but are closely linked are exceptionally short. A simple intuitive explanation for this pattern suggests itself - loci physically linked to the selected sites can only survive on the wrong side of the hybrid zone for a reasonable amount of time if they recombine away from their ancestry at the locally adaptive site. Because it is then tightly linked to the alternate ancestry at the selected site, this ancestry chunk will large recombine with the alternate ancestry, and therefore decay rapidly. This pattern means that we can further refine the signal of block lengths around a selected locus, by measuring the ratio $\frac{2\sum{l_B(m_i)}}{\sum{l_A(m_i+)+l_A(m_i-)}}$ of mean block lengths surrounding a site with the mean length of adjacent blocks (Figures~\ref{Supp:ratioBlockAdjacent},~\ref{Supp:ratioBlockAdjacentHeatmap})

@@

\begin{figure}
\includegraphics{figs/number_of_ancestors_tau1000}
\caption{The number of individuals of ancestry B per number of ancestors from $\tau=1000$ represented across geographic space. Each red or orange line represents a site some distance away (ranging from 0 - 0.01) from the selected site (here $s=0.01$ and $\sigma=1$). Yellow lines are corresponding positions on an unlinked chromosome with no selected loci, and grey lines are corresponding positions in a simulation with no selection.  Bold lines depict the target of selection when present, and corresponding position on chromosomes not harboring any selected sites.}\label{Fig:family_size}
\end{figure}
